Equalizer 2.0 --\\Convergence of a Parallel Rendering Framework

### 2018-02-14 paperIEEE.tex
Lots of little changes and a few last  fixmes.
git-svn-id: https://dev.ifi.uzh.ch/vmml/svn/Equalizer16@131 db7c848b-7755-4df2-b763-8796797afa52

\textsf{Collage} implements a few generic distributed objects which are used by
\textsf{Equalizer} and other applicatons. A barrier is a distributed barrier
primitive used for software swap barriers in \textsf{Equalizer}
(\sref{sec:swap}). Its implementation follows a simple master-slave approach,
which has shown to by sufficient for this use case. Queues are single producer,
multiple consumer FIFOs. To hide network latencies, consumers prefetch queue
items into a local queue. They are used for tile and chunk compounds
(\sref{sec:tile}).

@@

RTT Scale is used for a wide variety of use cases. In virtual reality, the
application is used for virtual prototyping and design reviews in front of
high-resolution display walls and CAVEs, as well as for virtual prototyping of
human-machine interactions CAVEs and HMDs (\fig{FIG_teaser}(b)). For
scalability, sort-first and tile compounds are used to achieve fast,
high-quality rendering, primarily for interactive raytracing, both based on CPUs
and GPUs. For CPU-based raytracing, often Linux-based rendering clients are used
with a Windows-based application node.

@@

RTNeuron \cite{HBBES:13} is a scalable real-time rendering tool for the
visualisation of neuronal simulations based on cable models. It is on
OpenSceneGraph for data management and Equalizer for parallel rendering, and
focuses not only on fast rendering times, but also on fast loading times with no
offline preprocessing. It provides level of detail (LOD) rendering, high quality
anti-aliasing based on jittered frusta and accumulation during still views,
interactive modification of the visual representation of neurons in a per-neuron
basis (full neuron vs. soma only, branch pruning depending on the branch level,
\dots). RTNeuron implements both sort-first and sort-last rendering with order
independent transparency.

@@

Omegalib \cite{Omegalib} is a software framework built on top of Equalizer that
facilitates application development for hybrid reality environments such as the
Cave 2. Hybrid reality enviroments aim to create a seamless 2D/3D environment
that supports both information-rich analysis (traditionally done on tiled
display wall) as well as virtual reality simulation exploration (traditionally
done in VR systems) at a resolution matching human visual acuity. Omegalib
supports dynamic reconfigurability of the display environment, so that areas of
the display can be interactively allocated to 2D or 3D workspaces as needed. It
makes it possible to have multiple immersive applications running on a
cluster-controlled display system, have different input sources dynamically
routed to applications, and have rendering results optionally redirected to a
distributed compositing manager. Omegalib supports pluggable front-ends, to
simplify the integration of third-party libraries like OpenGL, OpenSceneGraph,
and the Visualization Toolkit (VTK).

@@

The simpler \textsf{tree\_equalizer} outperforms in almost all cases the
load-grid-driven \textsf{load\_equalizer}, except for sort-first volume
rendering where the load in the region of interest is relatively uniform. This
counterintuitive result seems to again confirm that simple algorithms often
outperform theoretically better, but more complex implementations. On the other
hand, the \textsf{tile\_equalizer} often outperforms \textsf{tree\_equalizer}.
This suggests that the underlying implicit load balancing can be superior to the
explicit methods of \textsf{load\_equalizer} and \textsf{tree\_equalizer} in
high load situations, where the additional overhead of tile generation and
distribution is more justified. The relatively simple nature of our benchmark
application's rendering algorithms is also favoring work packages, since it has
a near-zero static overhead per rendering pass.

Last, but not least, we also provide scalability results for pixel compounds.
While naturally load-balanced, pixel compounds only scale fill rate and not
geometry processing. Consequently, pixel compounds provide better performance
for volume rendering, and a predictable scaling behaviour for both.

>>>>

\textsf{Collage} implements a few generic distributed objects which are used by
\textsf{Equalizer} and other applicatons. A barrier is a distributed barrier
primitive used for software swap barriers in \textsf{Equalizer}
(\sref{sec:swap}) \FIXME{Using the word barrier three times in one short sentence}. Its implementation follows a simple master-slave approach,
which has shown to by sufficient for this use case. The used distributed data structures are single producer,
multiple consumer FIFO queues. To hide network latencies, consumers prefetch stored
items into a local queue. Queues are used for tile and chunk compounds
(\sref{sec:tile}).

@@

RTT Scale is used for a wide variety of use cases. In virtual reality, the
application is used for virtual prototyping and design reviews in front of
high-resolution display walls and CAVEs, as well as for virtual prototyping of
human-machine interactions using CAVEs and HMDs (\fig{FIG_teaser}(b)). For
scalability, sort-first and tile compounds are used to achieve fast,
high-quality rendering, primarily for interactive raytracing, both based on CPUs
and GPUs. For CPU-based raytracing, often Linux-based rendering clients are used
with a Windows-based application node.

@@

RTNeuron \cite{HBBES:13} is a scalable real-time rendering tool for the
visualisation of neuronal simulations based on cable models. It is based on
OpenSceneGraph for data management and Equalizer for parallel rendering, and
focuses not only on fast rendering times, but also on fast loading times with no
offline preprocessing. It provides level of detail (LOD) rendering, high quality
anti-aliasing based on jittered frusta and accumulation during still views,
interactive modification of the visual representation of neurons on a per-neuron
basis (full neuron vs. soma only, branch pruning depending on the branch level,
\dots). RTNeuron implements both sort-first and sort-last rendering with order
independent transparency.

@@

Omegalib \cite{Omegalib} is a software framework built on top of Equalizer that
facilitates application development for hybrid reality environments such as the
Cave 2. Hybrid reality enviroments aim to create a seamless 2D/3D environment
that supports both information-rich analysis (traditionally done on tiled
display wall) as well as virtual reality simulation exploration (traditionally
done in VR systems) at a resolution matching human visual acuity. Omegalib
supports dynamic reconfigurability of the display environment, so that areas of
the display can be interactively allocated to 2D or 3D workspaces as needed. It
makes it possible to have multiple immersive applications running on a
cluster-controlled display system, have different input sources dynamically
routed to applications, and have rendering results optionally redirected to a
distributed compositing manager. Omegalib supports pluggable front-ends, to
simplify the integration of third-party libraries like OpenGL \FIXME{But if it's based on Equalizer then OpenGL should already be part of the backend?}, OpenSceneGraph,
and the Visualization Toolkit (VTK).

@@

The simpler \textsf{tree\_equalizer} outperforms in almost all cases the
load-grid-driven \textsf{load\_equalizer}, except for sort-first volume
rendering where the load in the region of interest is relatively uniform. This
counterintuitive result seems to again confirm that simple algorithms often
outperform theoretically better, but more complex implementations. On the other
hand, the \textsf{tile\_equalizer} often outperforms \textsf{tree\_equalizer}.
This suggests that the underlying implicit load balancing can be superior to the
explicit methods of \textsf{load\_equalizer} and \textsf{tree\_equalizer} in
high load situations, where the additional overhead of tile generation and
distribution is more justified. The relatively simple nature of our benchmark
application's rendering algorithms is also favoring work packages, since they have
a near-zero static overhead per rendering pass.

Finally, we also provide scalability results for pixel compounds.
While naturally load-balanced, pixel compounds only scale fill rate and not
geometry processing. Consequently, pixel compounds provide better performance
for volume rendering, and a predictable scaling behaviour for both.
